### 2024-11-22 paper/paper1.qmd

---
title: "Greatness"
subtitle: "Why Mahomes could be the GOAT"
author: 
  - Alexander Guarasci
thanks: "Code and data are available at: https://github.com/AlexanderG123/nfl"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---
It is widely accepted the Tom Brady is the greatest football player of all time
7 superbowls
etc
In recent years, and anecdotes that I have personally witnessed suggests that maybe there is an argument the mahomes could become the goat.

The estimand in this paper is whether or not Patrick Mahomes can usurp Thomas Edward Patrick Brady as the greatest football player of all time
```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

The data used in this paper was taken from nflverse [@citenflverse], which consists of “a set of packages dedicated to data of the National Football League.”

This analysis was conducted using the programming language R [@citeR] and supplemented with the analytical tools Tidyverse [@citetidyverse], Arrow [@citearrow] etc

The primary data utilized in this paper was QB data from every game throughout both Tom brady's career and Patrick Mahomes' career so far. The larger dataset consisted of 
## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

```{r}
#| label: fig-total-passing-yards
#| fig-cap: Total Passing Yards for Mahomes vs Brady
#| echo: false
#| warning: false
#| message: false

# Load and filter player game data
player_data <- read_parquet("~/nfl/data/analysis_data/game_stats/merged_player_game.parquet")
player_filtered <- player_data %>% 
  filter(player_display_name %in% c("Tom Brady", "Patrick Mahomes"))

# Total Passing Yards Plot
player_filtered %>%
  group_by(player_display_name) %>%
  summarise(
    total_passing_yards = sum(passing_yards, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = player_display_name, y = total_passing_yards, fill = player_display_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal() +
  labs(
    title = "Total Passing Yards: Mahomes vs Brady",
    x = "Player",
    y = "Passing Yards"
  )

```

```{r}
#| label: fig-total-rushing-yards
#| fig-cap: Total Rushing Yards for Mahomes vs Brady
#| echo: false
#| warning: false
#| message: false

# Total Rushing Yards Plot
player_filtered %>%
  group_by(player_display_name) %>%
  summarise(
    total_rushing_yards = sum(rushing_yards, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = player_display_name, y = total_rushing_yards, fill = player_display_name)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("blue", "red")) +
  theme_minimal() +
  labs(
    title = "Total Rushing Yards: Mahomes vs Brady",
    x = "Player",
    y = "Rushing Yards"
  )
```

```{r}
#| label: fig-passing-tds-per-season
#| fig-cap: Passing Touchdowns per Season for Mahomes and Brady
#| echo: false
#| warning: false
#| message: false

# Passing Touchdowns per Season Plot
player_data %>%
  group_by(player_display_name, season) %>%
  summarise(
    total_passing_tds = sum(passing_tds, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = season, y = total_passing_tds, color = player_display_name)) +
  geom_line() +
  theme_minimal() +
  labs(
    title = "Passing Touchdowns per Season",
    x = "Season",
    y = "Total Passing Touchdowns"
  )

```

```{r}
#| label: fig-passing-yard-distribution
#| fig-cap: Passing Yard Distribution for Mahomes and Brady
#| echo: false
#| warning: false
#| message: false

player_data %>%
  ggplot(aes(x = player_display_name, y = passing_yards, fill = player_display_name)) +
  geom_violin(alpha = 0.7,trim = FALSE) +
  theme_minimal() +
  scale_fill_manual(values = c("Mahomes" = "blue", "Brady" = "red")) +
  labs(
    title = "Passing Yards Distribution",
    x = "Player",
    y = "Passing Yards",
    fill = "Player"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14),
    legend.position = "none"
  )
```

```{r}
#| label: fig-interceptions-per-season
#| fig-cap: Interceptions per Season for Mahomes and Brady
#| echo: false
#| warning: false
#| message: false

# Interceptions per Season Plot
player_data %>%
  group_by(player_display_name, season) %>%
  summarise(
    total_interceptions = sum(interceptions, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = season, y = total_interceptions, color = player_display_name)) +
  geom_line() +
  theme_minimal() +
  labs(
    title = "Interceptions per Season",
    x = "Season",
    y = "Total Interceptions"
  )
```
Talk way more about it.

```{r}
#| label: fig-qbr-comparison
#| fig-cap: QBR Comparison by Season Type (Regular vs. Playoffs) for Brady and Mahomes
#| echo: false
#| warning: false
#| message: false

library(arrow)
library(ggplot2)
library(dplyr)

# Load and filter QBR data
qbr_data <- read_parquet("~/nfl/data/analysis_data/advanced_stats/merged_qbr.parquet")
qbr_filtered <- qbr_data %>% 
  filter(name_display %in% c("Tom Brady", "Patrick Mahomes"))

# QBR Comparison Plot
ggplot(qbr_filtered, aes(x = season_type, y = qbr_total, fill = name_display)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "QBR Comparison by Season Type (Regular vs. Playoffs)",
    x = "Season Type",
    y = "QBR"
  )
```

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

## Model Overview

We designed a predictive model to estimate Patrick Mahomes' cumulative career statistics over an additional 200 games. The model leverages historical game data from Tom Brady as a benchmark and Patrick Mahomes' existing career data to generate predictions for key performance metrics.

### Mathematical Notation

The model is a multivariate linear regression defined as:

$$
\hat{y}_i = \beta_0 + \sum_{j=1}^p \beta_j x_{ij} + \epsilon_i
$$

Where:

- $\hat{y}_i$: Predicted value of the response variable (e.g., total passing yards).  
- $\beta_0$: Intercept term.  
- $\beta_j$: Coefficients for predictors $x_{ij}$, representing the influence of the $j$-th feature on the response variable.  
- $x_{ij}$: Observed value of the $j$-th predictor for the $i$-th observation.  
- $\epsilon_i$: Residual error term, assumed to be normally distributed with mean 0.

We estimate $\beta_0$ and $\beta_j$ using Ordinary Least Squares (OLS) on Tom Brady's historical data, as his career provides a well-documented trajectory for long-term performance.

### Variables and Justification

The predictors ($x_j$) included in the model are chosen based on their relevance to a quarterback's performance:

1. **Passing Metrics**:  
   - `completions`, `attempts`, `passing_yards`, `passing_tds`, `interceptions`, `passing_air_yards`, `passing_first_downs`.  
   - These metrics directly influence a quarterback’s overall contribution to team success.

2. **Rushing Metrics**:  
   - `carries`, `rushing_yards`, `rushing_tds`, `rushing_first_downs`.  
   - While not the primary focus, rushing performance is crucial for dual-threat quarterbacks like Mahomes.

3. **Ball Security**:  
   - `sack_fumbles`, `rushing_fumbles`.  
   - Turnovers are critical in evaluating overall impact.

These features ensure the model captures both primary and auxiliary aspects of quarterback performance.

### Model Assumptions

1. **Linearity**: The relationship between predictors and response variables is linear.  
2. **Independence**: Residuals are independent across observations.  
3. **Homoscedasticity**: Residuals have constant variance.  
4. **Normality**: Residuals are normally distributed.  
5. **Stationarity**: Mahomes' performance trajectory will follow trends observed in Brady's career.

### Software and Implementation

The model was implemented using R, leveraging the `lm` function from the `stats` package for regression. The dataset was preprocessed using `tidyverse`, and predictions were calculated for an additional 200 games using the coefficients estimated from Brady's data.

### Validation and Diagnostics

1. **Train-Test Split**: The Brady dataset was split into training (80%) and testing (20%) sets. The model was trained on the training set and validated on the testing set.  
2. **Error Metrics**:  
   - **Root Mean Squared Error (RMSE)**: Measures the model's predictive accuracy.  
   - **Mean Absolute Error (MAE)**: Evaluates average prediction error.  
3. **Residual Analysis**:  
   - Residual plots confirmed the assumptions of homoscedasticity and normality.  
4. **Out-of-Sample Testing**: Predicted Mahomes' statistics on his observed data and compared with actual results to ensure alignment.

### Alternative Models Considered

1. **Decision Trees**:  
   - Strengths: Captures non-linear relationships.  
   - Weaknesses: Tends to overfit without pruning; less interpretable than linear regression.

2. **Bayesian Regression**:  
   - Strengths: Allows incorporation of priors, producing probabilistic predictions.  
   - Weaknesses: Increased complexity and computational requirements.

3. **Final Choice**:  
   - Linear regression was chosen for its balance of simplicity, interpretability, and performance. It aligns with the assumption that a quarterback’s performance trends over time can be captured linearly.

### Limitations

1. **Career Longevity**:  
   - The model assumes Mahomes will continue performing at a consistent level for 200 games. This may not account for potential injuries, performance decline, or external factors.  
2. **Sample Bias**:  
   - Relying on Brady's career as a benchmark may introduce bias, as it assumes Mahomes will follow a similar trajectory.  
3. **Feature Engineering**:  
   - Excluding contextual factors like team strength or play style may limit predictive accuracy.

### Conclusion

The linear regression model provides a robust yet interpretable framework for predicting Mahomes' lifetime stats. It is grounded in historical data and validated through residual analysis and error metrics. While limitations exist, the model effectively captures the essence of Mahomes' expected performance trajectory based on Brady's historical patterns.

\newpage

# References